---
title: "Lab 10 Student Problems"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


## Student Problem 01

We will be working with a cleaned version of last week's dataset that examines the relationship between employment type and the occurrence of stroke.

(1) Load the dataset into your environment and take a look.

```{r}

library(tidyverse)

stroke_raw <- read_csv("data/stroke_data.csv")

```

(2) Create a descriptive table of the entire dataset using gtsummary::tbl_summary().

Include better labels for each variable.

Change the summary statistic from Median(IQR) to Mean(SD); (Min, Max)

```{r}

library(gtsummary)

stroke_raw %>%
  select(stroke, everything(), -id) %>% #ordering the variables
  mutate(bmi = as.numeric(bmi)) %>% #fixing BMI to be numeric
  mutate(smoking_status = factor(smoking_status, levels = c("never smoked", #changing the order of the levels
                                                            "formerly smoked",
                                                            "smokes"))) %>% 
  tbl_summary(label = list(stroke ~ "Stroke Occurred",
                           gender ~ "Gender",
                           age ~ "Age",
                           hypertension ~ "Hypertension Positive",
                           heart_disease ~ "Heart Disease Positive",
                           Residence_type ~ "Residence Type",
                           ever_married ~ "Ever Married",
                           avg_glucose_level ~ "Average Glucose Level",
                           bmi ~ "BMI",
                           smoking_status ~ "Smoking Status",
                           work_cl ~ "Employment Type"),
              missing_text = "Missing", # change the text used to describe missing
              statistic = all_continuous() ~ "{mean} ({sd}); ({min}, {max})") %>% # custom summary statistics
  bold_labels()


```

(3) Create a new table split by employment category.

Add a column that shows the p-value for tests for differences in each variable across work categories.

Ensure that all tests work and that continuous variables are being tested using one-way ANOVA.

```{r}

stroke_raw %>%
  select(stroke, everything(), -id) %>%
  mutate(bmi = as.numeric(bmi)) %>%
  mutate(smoking_status = factor(smoking_status, levels = c("never smoked",
                                                            "formerly smoked",
                                                            "smokes"))) %>%
  tbl_summary(by = work_cl, # add by to create different columns split 
              label = list(stroke ~ "Stroke Occurred",
                           gender ~ "Gender",
                           age ~ "Age",
                           hypertension ~ "Hypertension Positive",
                           heart_disease ~ "Heart Disease Positive",
                           Residence_type ~ "Residence Type",
                           ever_married ~ "Ever Married",
                           avg_glucose_level ~ "Average Glucose Level",
                           bmi ~ "BMI",
                           smoking_status ~ "Smoking Status"),
              missing_text = "Missing",
              statistic = all_continuous() ~ "{mean} ({sd}); ({min}, {max})") %>%
  bold_labels() %>%
  add_overall() # add an overall column

# add a p-value

stroke_raw %>%
  select(stroke, everything(), -id) %>%
  mutate(bmi = as.numeric(bmi)) %>%
  mutate(smoking_status = factor(smoking_status, levels = c("never smoked",
                                                            "formerly smoked",
                                                            "smokes"))) %>%
  tbl_summary(by = work_cl, 
              label = list(stroke ~ "Stroke Occurred",
                           gender ~ "Gender",
                           age ~ "Age",
                           hypertension ~ "Hypertension Positive",
                           heart_disease ~ "Heart Disease Positive",
                           Residence_type ~ "Residence Type",
                           ever_married ~ "Ever Married",
                           avg_glucose_level ~ "Average Glucose Level",
                           bmi ~ "BMI",
                           smoking_status ~ "Smoking Status"),
              missing_text = "Missing",
              statistic = all_continuous() ~ "{mean} ({sd}); ({min}, {max})") %>%
  bold_labels() %>%
  add_p(test = all_continuous() ~ "aov", # adding a p-value
        test.args = all_tests("fisher.test") ~ list(simulate.p.value = TRUE)) # change options for a test that generates the pvalue


```

(4) Calculate the proportion positive for stroke and it's 95% confidence interval for each work category using the traditional Wald formula:

$$\hat{p}\pm1.96*\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}$$

```{r}

prop_summary <- stroke_raw %>%
  group_by(work_cl) %>%
  summarize(n_stroke = sum(stroke),
            n_total = n()) %>%
  mutate(p_hat = n_stroke/n_total,
         se = sqrt(p_hat*(1-p_hat)/n_total),
         lower_ci = p_hat - 1.96*se,
         upper_ci = p_hat + 1.96*se)

```

(5) Graph these confidence intervals using `geom_pointrange()`.

```{r}

ggplot(data = prop_summary) +
  geom_pointrange(aes(x = work_cl, y = p_hat, ymin = lower_ci, ymax = upper_ci)) +
  theme_bw()



```

(6) Now calculate the same confidence intervals using bootstrapping and create a faceted graph of the bootstrapped distributions by work category.


```{r}

# Function to obtain proportions
get_proportions <- function(df){
  prop_out <- df %>%
    group_by(work_cl) %>%
    summarize(n_stroke = sum(stroke),
              n_total = n()) %>%
    mutate(p_hat = n_stroke/n_total)
  return(prop_out)
}

#testing that it works for observed data
get_proportions(stroke_raw)

# Function to create a bootstrap sample
create_bootstrap <- function(df){
  bootstrap_out <- df %>%
    slice_sample(prop = 1, replace = TRUE)
  return(bootstrap_out)
}

# check that there is only one observation per id
stroke_raw %>%
  group_by(id) %>%
  tally() %>%
  arrange(desc(n)) %>% View()

# test the bootstrap creation
boot_ex <- create_bootstrap(stroke_raw)

# see that the bootstrapped sample works as intended, multiple rows per id
boot_ex %>%
  group_by(id) %>%
  tally() %>%
  arrange(desc(n)) %>% View()

# Function that gets a bootstrapped sample and then calculates proportion
get_boot_proportion <- function(df){
  out <- create_bootstrap(df) %>%
    get_proportions()
  return(out)
}

# compare our original summary to the summary of the bootstrapped data
prop_summary

get_boot_proportion(stroke_raw)

# Getting all the proportions together:
bootstrap_results <- map_dfr(1:2000, function(x) get_boot_proportion(stroke_raw))

# Plot bootstrapped distribution by work category
ggplot(data = bootstrap_results) +
  geom_histogram(aes(x = p_hat)) +
  facet_wrap(~work_cl, scales = "free") +
  theme_bw()

# Calculate the bootstrapped CI
bootstrapped_ci <- bootstrap_results %>%
  group_by(work_cl) %>%
  summarize(lower_ci = quantile(p_hat, 0.025),
            upper_ci = quantile(p_hat, 0.975))

```

(7) Compare your Wald and bootstrapped CI.

```{r}

bootstrapped_ci

prop_summary %>%
  select(work_cl, lower_ci, upper_ci)



```

They are pretty close!
