---
title: "Lab 10 Student Problems"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


## Student Problem 01

We will be working with a cleaned version of last week's dataset that examines the relationship between employment type and the occurrence of stroke.

(1) Load the dataset into your environment and take a look.

```{r}


```

(2) Create a descriptive table of the entire dataset using gtsummary::tbl_summary().

Include better labels for each variable.

Change the summary statistic from Median(IQR) to Mean(SD); (Min, Max)

```{r}



```

(3) Create a new table split by employment category.

Add a column that shows the p-value for tests for differences in each variable across work categories.

Ensure that all tests work and that continuous variables are being tested using one-way ANOVA.

```{r}



```

(4) Calculate the proportion positive for stroke and it's 95% confidence interval for each work category using the traditional Wald formula:

$$\hat{p}\pm1.96*\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}$$

```{r}



```

(5) Graph these confidence intervals using `geom_pointrange()`.

```{r}



```

(6) Now calculate the same confidence intervals using bootstrapping and create a faceted graph of the bootstrapped distributions by work category.


```{r}



```

(7) Compare your Wald and bootstrapped CI.
